# **CONTROL ANALYSIS — Fit Surrogate Models**

**Author:** Marcos E. Domínguez Arriola
**Repository:** https://github.com/elidom/EEG-Hyperscanning-Code_How-do-we-align-in-good-conversations-project

**Project:** Domínguez-Arriola, Lam, Pérez, & Pell (under review)
How Do We Align in Good Conversation?

Fits Linear Mixed-Effects Models to surrogate (shuffled-dyad) neural datasets to construct permutation-based null distributions for the relationship between perceived interaction quality (PIQ) and inter-brain relationships.

**Notes:**
- The user must manually specify:
  * frequency band (`alpha` or `theta`)
  * temporal dependency (`concurrent` or `recurrent`)


Libraries:
```{r}
library(tidyverse)
library(patchwork)
library(here)
library(lme4)
library(lmerTest)
library(emmeans)
```

Parameters
```{r}
freqband    <- "alpha"
temp_dep    <- "recurrent"
srr_samples <- 1:200

# optional verbose:
show_Fs <-
  FALSE # whether to print resulting PIQ mainf effects (F statistic) on each loop iteration

which_rois_to_print <-
  "right_temporal-medial_anterior" # which ROI pair to get live info on
```

Input directory
```{r}
if (temp_dep=="concurrent") {
  input_dir <- here("data", "9b_shuffled_surrogates_sync")
  
} else if (temp_dep == "recurrent") {
  input_dir <- here("data", "9b_shuffled_surrogates_rcrr")
  
} else {
  print("Error! Temporal dependency has to be 'concurrent' or 'recurrent'")
}
``` 

Helpers:
```{r}
source("functions/load_or_fit.R")

autoRefit <- function(mod) {
  convergedCheck = buildmer::converged(mod)
  
  attemptNum = 1
  
  while (convergedCheck == FALSE) {
    print(paste("Attempt number", attemptNum))
    fitVals = getME(mod, "theta")
    mod <- update(mod, control = strict_tol, start = fitVals)
    convergedCheck = buildmer::converged(mod)
    attemptNum = attemptNum + 1
  }
  
  return(mod)
  
}

options(scipen = 999, digits = 3)

rescale <- arm::rescale
lmer <- lmerTest::lmer
select <- dplyr::select

strict_tol <- lmerControl(optCtrl = list(xtol_abs = 1e-8, ftol_abs = 1e-8))

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::opts_chunk$set(dpi = 600)
```

Other preparations:
```{r}
electrodes <- c(
  "Fp1", "Fz", "F3", "F7", "FT9", "FC5", "FC1", "C3", "T7", "TP9", 
  "CP5", "CP1", "Pz", "P3", "P7", "O1", "Oz", "O2", "P4", "P8", 
  "TP10", "CP6", "CP2", "Cz", "C4", "T8", "FT10", "FC6", "FC2", 
  "F4", "F8", "Fp2", "FCz"
)

roi_mapping <- list(
  "left_anterior" = c("Fp1", "F3", "F7"),
  "right_anterior" = c("Fp2", "F4", "F8"),
  "medial_anterior" = c("Fz", "FC1", "FC2", "FCz"),
  "medial_central" = c("C3", "Cz", "C4", "CP1", "CP2"),
  "left_temporal" = c("FC5", "T7", "CP5"),
  "right_temporal" = c("FC6", "T8", "CP6")
)

roi_ord = c("left_anterior", "medial_anterior", "right_anterior", "medial_central", "left_temporal", "right_temporal")

# Reverse the mapping
electrode_to_roi <- unlist(lapply(names(roi_mapping), 
                                  function(x) setNames(rep(x, length(roi_mapping[[x]])),
                                                       roi_mapping[[x]])))

# Create a dataframe from the mapping
roi_df <- data.frame(chlabel = names(electrode_to_roi), 
                     roi = electrode_to_roi, 
                     stringsAsFactors = FALSE
) 

```


```{r}
all_Fs <- tibble()
all_emt <- tibble()

start.time <- Sys.time()

for (i in srr_samples) {
  df <-
    readRDS(
      paste0(
        input_dir,
        "/Shuffled_GCMI_",
        str_to_sentence(freqband),
        "_",
        as.character(i),
        ".RDS"
      )
    )
  
  names(df)[names(df) %in% c("SyncMI", "RcrrMI")] <- "MI"
  
  df_tidy <-
    df %>%
    left_join(roi_df, by = c("Channel_A" = "chlabel")) %>%
    rename(ROI_A = roi) %>%
    left_join(roi_df, by = c("Channel_B" = "chlabel")) %>%
    rename(ROI_B = roi) %>%
    filter(!is.na(ROI_A) & !is.na(ROI_B)) %>%
    
    mutate(ROI_pair = paste(ROI_A, ROI_B, sep = "-")) %>%
    
    mutate(
      ROI_B = factor(ROI_B, levels = roi_ord),
      ROI_A = factor(ROI_A, levels = roi_ord)
    ) %>%
    mutate(dplyr::across(
      c(Channel_A, Channel_B,
        ROI_A, ROI_B, ROI_pair),
      ~ as.factor(.)
    )) %>%
    
    mutate(elec_pair = factor(paste(Channel_A, Channel_B, sep = "_"))) %>%
    
    # square root-transform GCMI
    mutate(
      MI = pmax(MI, 0),
      sq_MI = sqrt(MI),
      SurDyad   = factor(paste(DyadA, DyadB, sep = "-")),
      TrialID = factor(paste(Type, block, SurDyad, sep = "_"))
    ) %>%
    select(Band,
           SurDyad,
           Speaker,
           Type,
           block,
           PIQ_mean,
           sq_MI,
           MI,
           elec_pair,
           ROI_pair,
           TrialID)
  
  df_scaled <- df_tidy %>%
    mutate(sq_MI = as.numeric(scale(sq_MI)),
           PIQ_mean  = as.numeric(scale(PIQ_mean)))
  
  
  # Model Fit (Same as real analysis)
  mod <- load_or_fit(
    model_name = paste0("surr_model_", freqband, "_", temp_dep, "_" , as.character(i)),
    models_base_dir = "surr_models",
    modelSpec =
      "lmer(sq_MI ~ PIQ_mean * ROI_pair +
                         (1 | SurDyad) + (1 + PIQ_mean | SurDyad : Speaker) +
                         (1 | elec_pair),
                       data = df_scaled, REML = F)",
    max_refit = 10
  )
  
  
  anova_results <- anova(mod)
  
  PIQ_F  <- anova_results$`F value`[1]
  PIQ_df <- tibble(Fval = PIQ_F,
                   surr = i)
  
  emtrends_mod <- emtrends(
    mod,
    ~ ROI_pair,
    var = "PIQ_mean",
    lmerTest.limit = 254016,
    lmer.df = "satterthwaite"
  )
  
  em_results <- summary(
    emtrends_mod,
    adjust = "fdr",
    infer = c(TRUE, TRUE),
    level = 0.95
  )
  
  
  em_df <- tibble(
    ROI_pair = em_results$ROI_pair,
    coef     = em_results$PIQ_mean.trend,
    t        = em_results$t.ratio,
    surr     = i
  )
  
  
  all_emt <- rbind(all_emt, em_df)
  all_Fs <- rbind(all_Fs, PIQ_df)
  
  passed.time <- Sys.time()
  time.taken <- passed.time - start.time
  
  print(paste0(
    "Surrogate ",
    as.character(i),
    " completed. t = ",
    as.character(round(em_df[em_df$ROI_pair == which_rois_to_print, ]$t, 2))
  ))
  
  if (show_Fs) {
    print(paste0("F = ", as.character(round(PIQ_F, 2))))
  }
  
  print(time.taken)
}

```

```{r}
saveRDS(
  all_emt,
  paste0(
    "data/10_surrogate_outputs/",
    freqband,
    "_",
    temp_dep,
    "_",
    "surrogate_emts.rds"
  )
)

saveRDS(
  all_Fs,
  paste0(
    "data/10_surrogate_outputs/",
    freqband,
    "_",
    temp_dep,
    "_",
    "surrogate_Fs.rds"
  )
)
```

